---
title: "A cursory exploration of trends in National Health Expenditure in Canada from 1985 to 2025"
output: 
  html_document:
    toc: TRUE
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE)

library(easypackages)

libraries(
  c(
    "ggplot2", 
    "tidyr", 
    "tidyverse", 
    "ddply", 
    "ggpubr", 
    "janitor", 
    "stringr", 
    "readxl", 
    "scales"
  )
)
```

This page presents a curated set of descriptive visualizations of healthcare spending by sector. Specific expenditures include hospitals, physicians, medications, public health, administration, COVID-19 response and other professionals.

```{r}
dataA_00 = read_xlsx(
  path = "ANALYSIS-nhex-open-data-2025-en.xlsx", 
  sheet = "Table O.1"
) %>%
  clean_names() %>%
  mutate(
    province = 
      factor(
        province, 
        levels = 
          c("Alberta", "British Columbia", "Manitoba", "New Brunswick", "Newfoundland and Labrador", "Northwest Territories", "Nova Scotia", "Nunavut", "Ontario", "Prince Edward Island", "Quebec", "Saskatchewan", "Yukon", "Canada")
        ), 
    sector = 
      factor(
        sector, levels = c("Private", "Provincial Government", "Territorial Government", "Public")
      ), 
    
    use_of_funds = 
      factor(
        use_of_funds
      ), 
    
    current_dollars = current_dollars %>% as.numeric(), 
    
    current_dollars_per_capita = current_dollars_per_capita %>% as.numeric()
  )


# dataA_00$year %>% table(useNA = "always")
# class(dataA_00$year)
# 
# 
# dataA$forecast_category %>% table(useNA = "always")
# 
# dataA_00$province %>% table(useNA = "always")
# class(dataA_00$province)
# 
# dataA_00$sector %>% table(useNA = "always")
# 
# dataA_00$use_of_funds %>% table(useNA = "always")
# 
# dataA_00$current_dollars %>% table(useNA = "always")
# 
# dataA_00$current_dollars_per_capita %>% class
# 
# dataA_00$constant_2010_dollars


dataA = dataA_00 %>%
  filter(
    !province %in% c("Yukon", "Northwest Territories", "Nunavut")
  )








dataB_00 = 
  read_xlsx(path = "ANALYSIS-CA population in thousands.xlsx") %>%
  clean_names() %>%
  mutate(
    nun = nun %>% as.numeric
  )

dataB_01 = dataB_00 %>%
  pivot_longer(
    cols = n_l:canada, 
    names_to = "province", 
    values_to = "population"
  ) %>%
  group_by(
    province
  ) %>%
  mutate(
    diff_pop = population - lag(population), 
    year = year %>% as.numeric()
  ) %>%
  ungroup() %>%
  mutate(
    province = 
      plyr::mapvalues(
        province, 
        from = 
          c(
            "alta", 
            "b_c", 
            "man", 
            "n_b", 
            "n_l", 
            "n_s", 
            "n_w_t", 
            "ont", 
            "p_e_i",
            "y_t",
            "que", 
            "sask", 
            "nun"
          ), 
        to = 
          c(
            "Alberta", 
            "British Columbia", 
            "Manitoba", 
            "New Brunswick", 
            "Newfoundland and Labrador", 
            "Nova Scotia", 
            "Northwest Territories", 
            "Ontario", 
            "Prince Edward Island", 
            "Yukon",
            "Quebec", 
            "Saskatchewan", 
            "Nunavut"
          )
      )
  )


dataB = dataB_01

dataA$sector %>% unique
dataA$use_of_funds %>% unique

```

# Provincial Government health expenditure {.tabset}

## Total expenditure

```{r provincial gov total health expenditure, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "Total"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Hospitals

```{r provincial gov spending on hospitals, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "Hospitals"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Physicians

```{r provincial gov spending on physicians, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "Physicians"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Medications

```{r provincial gov spending on medications, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "Drugs"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Public Health

```{r provincial gov spending on public health, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "Public Health"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Administration

```{r provincial gov spending on admin, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "Administration"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On COVID-19 Response Funding

```{r provincial gov spending on covid, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "COVID-19 Response Funding"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Other Professionals

```{r provincial gov spending on other professionals, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector ="Provincial Government"
in_use_of_funds = "Other Professionals"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

# Health expenditure by public sector {.tabset}

## Total expenditure

```{r public total health expenditure, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "Total"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Hospitals

```{r public spending on hospitals, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "Hospitals"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Physicians

```{r public spending on physicians, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "Physicians"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Medications

```{r public spending on medications, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "Drugs"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Public Health

```{r public spending on public health, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "Public Health"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Administration

```{r public spending on admin, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "Administration"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On COVID-19 Response Funding

```{r public spending on covid, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "COVID-19 Response Funding"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Other Professionals

```{r public spending on other professionals, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Public"
in_use_of_funds = "Other Professionals"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

# Health expenditure by private sector {.tabset}

## Total expenditure

```{r private total health expenditure, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "Total"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Hospitals

```{r private spending on hospitals, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "Hospitals"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Physicians

```{r private spending on physicians, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "Physicians"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Medications

```{r private spending on medications, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "Drugs"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Public Health

```{r private spending on public health, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "Public Health"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Administration

```{r private spending on admin, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "Administration"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On COVID-19 Response Funding

```{r private spending on covid, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "COVID-19 Response Funding"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```

## On Other Professionals

```{r private spending on other professionals, include = TRUE, fig.width = 15, fig.height = 8}
# Custom formatting function to display "1 x 10^n"
scientific_10 <- function(x) {
  parse(text = gsub("e", " %*% 10^", scales::scientific_format()(x)))
}



in_sector = "Private"
in_use_of_funds = "Other Professionals"

p1 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    labels = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Total expenditure", width = 20), 
    title = "Total"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))

p2 = ggplot(
  dataA %>% filter(sector == in_sector, use_of_funds == in_use_of_funds, province != "Canada"), 
  aes(
    x = year, 
    y = current_dollars_per_capita
  )
) + 
  geom_point() +
  geom_line() +
  # geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    # limits = c(0, 700),
    # breaks = seq(0, 1000, 100),
    label = scientific_10
  ) +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x) 10^x), # Define the breaks as powers of 10
  #   labels = trans_format("log10", math_format(10^.x))  # Format labels as 10^exponent
  # ) +
  labs(
    x = "Year", 
    y = "Per capita expenditure", 
    title = "Per capita"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) +
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))








p3 = ggplot(
  dataB %>% filter(!province %in% c("Nunavut", "Northwest Territories", "canada", "Yukon") &  year != 2025 ), 
  aes(
    x = year, 
    y = diff_pop
  )
) + 
  geom_line(group = 1) +
  geom_vline(xintercept = 2020, linetype = "dotted") +
  geom_point() +
  # geom_smooth(method = "loess", se = FALSE, color = "black") +
  scale_x_continuous(
    limits = c(1985, 2029),
    breaks = seq(1905, 2030, 20)
  ) +
  scale_y_continuous(
    label = scientific_10
  ) +
  labs(
    x = "Year", 
    y = str_wrap("Population change from prior year (thousands)", width = 20), 
    title = "Change in population from previous year"
  ) +
  theme(
    text = element_text(size = 16, face = "bold"),
    plot.title = element_blank(), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.background = element_rect(fill = NA, color = "black"),
    panel.border = element_rect(fill = NA, color = "black"), 
    strip.background = element_rect(fill = NA, color = "black")
  ) + 
  facet_wrap(~province, nrow = 1, labeller = label_wrap_gen(width = 15))


ggarrange(p1, p2, p3, ncol = 1, align = "v") %>%
  annotate_figure(
    # top = 
    #   text_grob(
    #     paste0("Expenditure by ", in_sector, " for ", in_use_of_funds, " in current dollars"), 
    #     size = 18, 
    #     face = "bold"
    #     ),
      
    bottom = 
      text_grob(
        "Year", 
        size = 14, 
        face = "bold"
      )
  ) %>%
  annotate_figure(
    bottom = 
      text_grob(
        "Dotted vertical lines indicate year 2020.", 
        size = 12, 
        face = "bold", 
        hjust = 2
      )
  )
```
