---
title: "Overview of tidymodels"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidymodels)
library(dials)
library(DiagrammeR)
set.seed(123)
```

# End-to-end overview of building a machine learning model

1.  Split data in training and test set.
    -   split = initial_split(data, prop); train = training(split), test = testing(split)
2.  Preprocess recipe.
    -   rec = recipe(formula, data = train) %\>% step_novel(all_nominal_predictors()) %\>% step_dummy(all_nominal_predictors) %\>% step_zv(all_predictors()) %\>% step_normalize(all_continuous_predictors())
3.  Specify model.
    -   rf_spec = rand_forest(mtry = tune(), trees = tune()) %\>% set_engine(engine = "ranger") %\>% set_mode("classification)
4.  Specify workflow.
    -   wf = workflow() %\>% add_recipe(rec) %\>% add_model(rf_spec)
5.  Cross-validation/resampling scheme design.
    -   folds = vfold_cv(data = train, v = 5, strata = outcome)
6.  Tune hyperparameters.
    -   rf_parameters = parameters(mtry(range = c(3, 10)), trees(range = c(100, 5000)))
    -   rf_res = tune_grid(wf, grid = rf_parameters, resamples = folds, metrics = metric_set(accuracy, roc_auc))
    -   collect_metrics(rf_res); show_best(rf_res, metric = "roc_auc", n = 10)
7.  Finalize workflow with best parameters.
    -   best_rf = select_best(rf_res, metric = "roc_auc")
    -   rf_final_wf = finalize_workflow(rf_wf, best_rf)
8.  Final test set evaluation.
    -   rf_last = last_fit(rf_final_wf, split = split, metrics = metric_set(roc_auc, accuracy, sens, spec))
9.  Fit final model on training data (for interpretation, e.g. variable importance).
    -   rf_fit = fit(rf_final_wf, data = train)
    -   Variable importance: rf_engine = extract_fit_parsnip(rf_fit)\$fit; vip::vip(rf_engine, num_features)

# Templates

## Random forest template

```{r, eval = FALSE, include = TRUE, echo = TRUE}
# ============================================================
# Random Forest (tidymodels): End-to-End Workflow
# Using `parameters()` + grid generation (no expand.grid)
# ============================================================


# ------------------------------------------------------------
# Assumptions
# ------------------------------------------------------------
# df      : your data frame
# outcome : factor outcome variable for classification
# Event   : second factor level is treated as the "event" by yardstick

# ------------------------------------------------------------
# 1. Train / Test Split
# ------------------------------------------------------------
split <- initial_split(df, prop = 0.80, strata = outcome)
train <- training(split)
test  <- testing(split)

# ------------------------------------------------------------
# 2. Preprocessing Recipe
# ------------------------------------------------------------
rf_rec <- recipe(outcome ~ ., data = train) %>%
  step_novel(all_nominal_predictors()) %>%     # handle unseen levels
  step_dummy(all_nominal_predictors()) %>%     # one-hot encode factors
  step_zv(all_predictors())                    # remove zero-variance predictors

# ------------------------------------------------------------
# 3. Random Forest Model Spec (tunable params declared via tune())
# ------------------------------------------------------------
rf_spec <- rand_forest(
  mtry  = tune(),
  min_n = tune(),
  trees = tune()
) %>%
  set_engine("ranger", importance = "permutation") %>%
  set_mode("classification")

# ------------------------------------------------------------
# 4. Workflow
# ------------------------------------------------------------
rf_wf <- workflow() %>%
  add_recipe(rf_rec) %>%
  add_model(rf_spec)

# ------------------------------------------------------------
# 5. Cross-Validation
# ------------------------------------------------------------
folds <- vfold_cv(train, v = 5, strata = outcome)

# ------------------------------------------------------------
# 6. Define parameter ranges with `parameters()`
# ------------------------------------------------------------
# NOTE: mtry() is data-dependent (post-recipe predictor count), so tidymodels may
# finalize it after prepping the recipe. You can optionally set a range too.
rf_params <- parameters(
  mtry(range = c(2, 30)),          # adjust upper bound if you expect many dummies
  min_n(range = c(2, 40)),
  trees(range = c(500, 2000))
)

# Build a grid from those parameter definitions
# Option A: random grid (often better for 3+ params)
rf_grid <- grid_random(rf_params, size = 25)

# Option B: regular grid (can explode quickly: levels^(#params))
# rf_grid <- grid_regular(rf_params, levels = 5)

# ------------------------------------------------------------
# 7. Hyperparameter Tuning
# ------------------------------------------------------------
rf_res <- tune_grid(
  rf_wf,
  resamples = folds,
  grid = rf_grid,
  metrics = metric_set(roc_auc, accuracy),
  control = control_grid(save_pred = TRUE)
)

# Examine tuning results
collect_metrics(rf_res)
show_best(rf_res, metric = "roc_auc", n = 10)

# ------------------------------------------------------------
# 8. Finalize Workflow with Best Parameters
# ------------------------------------------------------------
best_rf <- select_best(rf_res, metric = "roc_auc")
rf_final_wf <- finalize_workflow(rf_wf, best_rf)

# ------------------------------------------------------------
# 9. Final Test-Set Evaluation (honest performance)
# ------------------------------------------------------------
rf_last <- last_fit(
  rf_final_wf,
  split = split,
  metrics = metric_set(roc_auc, accuracy, sens, spec)
)

collect_metrics(rf_last)

rf_test_preds <- collect_predictions(rf_last)

# ROC curve
rf_test_preds %>%
  roc_curve(outcome, .pred_2) %>%   # replace .pred_2 with your event class column if needed
  autoplot()

# ------------------------------------------------------------
# 10. Fit Final Model on Training Data (for interpretation)
# ------------------------------------------------------------
rf_fit <- fit(rf_final_wf, data = train)

# Variable importance (since ranger importance = "permutation")
rf_engine <- extract_fit_parsnip(rf_fit)$fit
vip::vip(rf_engine, num_features = 20)

# ------------------------------------------------------------
# (Optional) Partial Dependence Plot via DALEX (workflow-safe)
# ------------------------------------------------------------
library(DALEX)
library(DALEXtra)

explainer <- explain_tidymodels(
  rf_fit,
  data = train,
  y = train$outcome,
  verbose = FALSE
)

model_profile(explainer, variables = "age") %>% plot()

```

## kNN template

```{r, eval = FALSE, include = TRUE, echo = TRUE}
# ============================================================
# k-Nearest Neighbors (tidymodels): End-to-End Workflow
# ============================================================

library(tidymodels)
set.seed(123)

# ------------------------------------------------------------
# Assumptions
# ------------------------------------------------------------
# df        : your data frame
# outcome   : factor outcome variable for classification
# NOTE: k-NN is distance-based -> normalization is essential.

# ------------------------------------------------------------
# 1. Train / Test Split
# ------------------------------------------------------------
split <- initial_split(df, prop = 0.80, strata = outcome)
train <- training(split)
test  <- testing(split)

# ------------------------------------------------------------
# 2. Preprocessing Recipe (important for k-NN!)
# ------------------------------------------------------------
knn_rec <- recipe(outcome ~ ., data = train) %>%
  step_novel(all_nominal_predictors()) %>%     # handle unseen levels
  step_dummy(all_nominal_predictors()) %>%     # one-hot encode factors
  step_zv(all_predictors()) %>%                # remove zero-variance predictors
  step_normalize(all_numeric_predictors())     # SCALE for distances

# ------------------------------------------------------------
# 3. k-NN Model Specification (with tuning)
# ------------------------------------------------------------
knn_spec <- nearest_neighbor(
  neighbors   = tune(),     # k
  dist_power  = tune(),     # 1 = Manhattan, 2 = Euclidean
  weight_func = tune()      # how neighbors are weighted
) %>%
  set_engine("kknn") %>%
  set_mode("classification")

# ------------------------------------------------------------
# 4. Workflow
# ------------------------------------------------------------
knn_wf <- workflow() %>%
  add_recipe(knn_rec) %>%
  add_model(knn_spec)

# ------------------------------------------------------------
# 5. Cross-Validation
# ------------------------------------------------------------
folds <- vfold_cv(train, v = 5, strata = outcome)

# ------------------------------------------------------------
# 6. Parameter Ranges + Grid
# ------------------------------------------------------------
knn_params <- parameters(
  neighbors(range = c(3, 35)),
  dist_power(range = c(1, 2)),
  weight_func(values = c("rectangular", "triangular", "gaussian"))
)

knn_grid <- grid_regular(knn_params, levels = 5)
# If this is too big (levels^#params), use a random grid instead:
# knn_grid <- grid_random(knn_params, size = 30)

# ------------------------------------------------------------
# 7. Hyperparameter Tuning
# ------------------------------------------------------------
knn_res <- tune_grid(
  knn_wf,
  resamples = folds,
  grid = knn_grid,
  metrics = metric_set(roc_auc, accuracy),
  control = control_grid(save_pred = TRUE)
)

# Examine tuning results
collect_metrics(knn_res)
show_best(knn_res, metric = "roc_auc", n = 10)

# ------------------------------------------------------------
# 8. Finalize Workflow with Best Parameters
# ------------------------------------------------------------
best_knn <- select_best(knn_res, metric = "roc_auc")

knn_final_wf <- finalize_workflow(knn_wf, best_knn)

# ------------------------------------------------------------
# 9. Final Test-Set Evaluation (honest performance)
# ------------------------------------------------------------
knn_last <- last_fit(
  knn_final_wf,
  split = split,
  metrics = metric_set(roc_auc, accuracy, sens, spec)
)

collect_metrics(knn_last)

knn_test_preds <- collect_predictions(knn_last)

# ROC curve
knn_test_preds %>%
  roc_curve(outcome, .pred_2) %>%   # replace .pred_2 with your event class column if needed
  autoplot()

# ------------------------------------------------------------
# 10. Fit Final Model on Training Data (for interpretation/usage)
# ------------------------------------------------------------
knn_fit <- fit(knn_final_wf, data = train)

# (Optional) Confusion matrix at threshold 0.5
knn_test_preds %>%
  mutate(.pred_class = factor(if_else(.pred_2 >= 0.5, levels(train$outcome)[2], levels(train$outcome)[1]),
                              levels = levels(train$outcome))) %>%
  conf_mat(outcome, .pred_class)

```

## Neural networks

```{r, eval = FALSE, include = TRUE, echo = TRUE}
# ============================================================
# Neural Net (tidymodels): End-to-End Workflow
# Using `mlp()` + `parameters()` + tuning + test evaluation
# Engine: nnet (lightweight, good for portfolios)
# ============================================================

library(tidymodels)
library(dials)
library(rlang)

set.seed(123)

# ------------------------------------------------------------
# Assumptions
# ------------------------------------------------------------
# df      : your data frame
# outcome : factor outcome variable for classification
# NOTE: Neural nets are scale-sensitive -> normalization is essential.

# ------------------------------------------------------------
# 1. Train / Test Split
# ------------------------------------------------------------
split <- initial_split(df, prop = 0.80, strata = outcome)
train <- training(split)
test  <- testing(split)

# Identify the "event" class (tidymodels/yardstick default is the 2nd level)
event_level <- levels(train$outcome)[2]
prob_col <- paste0(".pred_", event_level)

# ------------------------------------------------------------
# 2. Preprocessing Recipe (important for neural nets!)
# ------------------------------------------------------------
nn_rec <- recipe(outcome ~ ., data = train) %>%
  step_novel(all_nominal_predictors()) %>%      # handle unseen levels
  step_dummy(all_nominal_predictors()) %>%      # one-hot encode factors
  step_zv(all_predictors()) %>%                 # remove zero-variance predictors
  step_normalize(all_numeric_predictors())      # SCALE numeric predictors

# ------------------------------------------------------------
# 3. Neural Net Model Spec (tunable params declared via tune())
# ------------------------------------------------------------
# `mlp()` = multilayer perceptron
# hidden_units: size of hidden layer
# penalty     : weight decay (regularization)
# epochs      : training iterations
nn_spec <- mlp(
  hidden_units = tune(),
  penalty      = tune(),
  epochs       = tune()
) %>%
  set_engine("nnet") %>%
  set_mode("classification")

# ------------------------------------------------------------
# 4. Workflow
# ------------------------------------------------------------
nn_wf <- workflow() %>%
  add_recipe(nn_rec) %>%
  add_model(nn_spec)

# ------------------------------------------------------------
# 5. Cross-Validation
# ------------------------------------------------------------
folds <- vfold_cv(train, v = 5, strata = outcome)

# ------------------------------------------------------------
# 6. Define parameter ranges with `parameters()`
# ------------------------------------------------------------
# Notes:
# - penalty() is on a log10 scale internally (dials handles that)
# - epochs can be modest; too high can overfit and waste time
nn_params <- parameters(
  hidden_units(range = c(1L, 50L)),
  penalty(range = c(-6, -1)),          # 10^-6 to 10^-1 (log10 scale)
  epochs(range = c(50L, 400L))
)

# Random grid is usually better than regular for 3 parameters
nn_grid <- grid_random(nn_params, size = 25)

# ------------------------------------------------------------
# 7. Hyperparameter Tuning
# ------------------------------------------------------------
nn_res <- tune_grid(
  nn_wf,
  resamples = folds,
  grid = nn_grid,
  metrics = metric_set(roc_auc, accuracy),
  control = control_grid(save_pred = TRUE)
)

# Examine tuning results
collect_metrics(nn_res)
show_best(nn_res, metric = "roc_auc", n = 10)

# ------------------------------------------------------------
# 8. Finalize Workflow with Best Parameters
# ------------------------------------------------------------
best_nn <- select_best(nn_res, metric = "roc_auc")
nn_final_wf <- finalize_workflow(nn_wf, best_nn)

# ------------------------------------------------------------
# 9. Final Test-Set Evaluation (honest performance)
# ------------------------------------------------------------
nn_last <- last_fit(
  nn_final_wf,
  split = split,
  metrics = metric_set(roc_auc, accuracy, sens, spec)
)

collect_metrics(nn_last)

nn_test_preds <- collect_predictions(nn_last)

# ROC curve (uses the event class probability column dynamically)
nn_test_preds %>%
  roc_curve(outcome, !!sym(prob_col)) %>%
  autoplot()

# Optional: confusion matrix at 0.5 threshold (you can choose a better threshold later)
nn_test_preds %>%
  mutate(
    .pred_class = factor(
      if_else(!!sym(prob_col) >= 0.5, event_level, levels(train$outcome)[1]),
      levels = levels(train$outcome)
    )
  ) %>%
  conf_mat(outcome, .pred_class)

# ------------------------------------------------------------
# 10. Fit Final Model on Training Data (for interpretation/usage)
# ------------------------------------------------------------
nn_fit <- fit(nn_final_wf, data = train)

# Inspect finalized model specification
extract_spec_parsnip(nn_fit)

```

# Reporting Machine Learning Results: Practical Guidelines

Below is a compilation of advice for a practical <!---, reviewer- and hiring-manager-friendly--->framework for reporting machine learning results, with random forest used as a concrete example. <!---This applies equally to RPubs, papers, internal reports, and portfolios.--->

The structure is as follows: - Each section header describes what to report - A discussion of common mistakes - Example text using random forest - Key points people are usually looking for

------------------------------------------------------------------------

# A general template for reporting ML results

Think in terms of questions the reader has.

------------------------------------------------------------------------

## A. Objective & prediction task (always first)

**Answer:**\
What problem is the model solving?

**Include:** - Outcome definition - Population - Prediction vs explanation - Unit of prediction (patient, visit, image, etc.)

**Example:**\
“We developed a machine learning model to predict 30-day mortality among hospitalized patients using routinely collected clinical variables.”

🚫 **Don’t start with algorithms. Start with the problem.**

------------------------------------------------------------------------

## B. Data & study design

**Answer:**\
Where did the data come from, and how were they used?

**Include:** - Sample size - Outcome prevalence - Train/test split - Cross-validation strategy - Any stratification or grouping

**Example:**\
“The dataset consisted of 4,812 patients, of whom 9.6% experienced the outcome. Data were split into training (80%) and test (20%) sets, stratified by outcome. Hyperparameters were optimized using 5-fold cross-validation on the training data.”

This signals methodological literacy.

------------------------------------------------------------------------

## C. Preprocessing (this matters more than people think)

**Answer:**\
How was the data prepared, and was leakage avoided?

**Include:** - Handling of categorical variables - Scaling (if applicable) - Missingness handling - Zero-variance filtering - Where preprocessing occurred (inside CV!)

**Example:**\
“Categorical predictors were dummy-encoded, and predictors with zero variance were removed using a recipe-based preprocessing pipeline. All preprocessing steps were estimated within resampling folds to prevent information leakage.”

That last sentence is very important!

------------------------------------------------------------------------

## D. Model specification & tuning

**Answer:**\
What model was fit, and how were hyperparameters chosen?

**Include:** - Algorithm - Engine/library - Tuned parameters - Optimization metric

**Random forest example:**\
“A random forest classifier was implemented using the ranger engine. The number of candidate predictors at each split (mtry) and the minimum node size (min_n) were optimized via grid search, selecting hyperparameters that maximized cross-validated ROC AUC.”

🚫 **Don’t just say “we fit a random forest.”**

------------------------------------------------------------------------

## E. Performance evaluation (this is the core)

**Answer:**\
How well does the model perform on unseen data?

**Always include:** - Held-out test set results - Primary metric - Secondary metrics - Confidence intervals if possible

**Example:**\
“On the held-out test set, the model achieved an ROC AUC of 0.82, with sensitivity of 0.74 and specificity of 0.79 at a probability threshold of 0.5.”

**Optional (but impressive):** - Calibration - Decision curves - Class imbalance handling

------------------------------------------------------------------------

## F. Model interpretation (with restraint)

**Answer:**\
How does the model behave, and what drives predictions?

**Include:** - Variable importance (with caveats) - Partial dependence / ALE / SHAP (if used) - Clear disclaimer about causality

**Example:**\
“Permutation-based variable importance indicated that age, baseline creatinine, and oxygen saturation contributed most strongly to predictive performance. Partial dependence plots suggested nonlinear risk increases at advanced ages; however, these plots reflect model behavior rather than causal effects.”

This shows statistical maturity.

------------------------------------------------------------------------

## G. Limitations (this builds trust)

**Answer:**\
What could go wrong or limit generalization?

**Always include:** - Observational data limitations - External validation status - Model interpretability limits

**Example:**\
“This analysis used retrospective single-source data, and external validation was not performed. Variable importance reflects predictive contribution rather than causal relevance.”

Reviewers expect this.

------------------------------------------------------------------------

## H. Practical implications / next steps

**Answer:**\
What would you do next if this mattered?

**Example:**\
“Future work includes external validation, calibration assessment, and comparison with simpler baseline models prior to potential deployment.”

This signals real-world thinking.

------------------------------------------------------------------------

# What NOT to do (very common mistakes)

🚫 Reporting CV performance as final results\
🚫 Claiming “important variables” are “risk factors”\
🚫 Omitting preprocessing details\
🚫 Showing only accuracy for imbalanced outcomes\
🚫 Treating ML outputs as causal evidence

Avoid these and you already beat many submissions.

------------------------------------------------------------------------

# Full example: concise random forest results section

Here’s a polished, copy-ready example:

> “A random forest classifier was developed to predict 30-day mortality using routinely collected clinical predictors. Data were split into training (80%) and test (20%) sets, stratified by outcome prevalence. Hyperparameters were optimized using 5-fold cross-validation on the training data. Preprocessing included dummy encoding of categorical variables and removal of zero-variance predictors within resampling folds to prevent data leakage.
>
> On the held-out test set, the final model achieved an ROC AUC of 0.82, with sensitivity of 0.74 and specificity of 0.79. Permutation-based variable importance identified age, baseline creatinine, and oxygen saturation as key contributors to predictive performance. Partial dependence plots demonstrated nonlinear associations between age and predicted risk; however, these plots describe model behavior and do not imply causal effects.
>
> This analysis is limited by the use of retrospective single-source data and the absence of external validation.”

------------------------------------------------------------------------

# What readers look for

When someone reads your report, they want to know if you've addressed the following: - Data leakage - Tuning vs evaluation: make sure these are separate events - Understanding the predictive abilities without overreaching on causal inference - Accessibility to non- or adjacent experts - Generalizability beyond the current dataset

------------------------------------------------------------------------

# TLDR cheat sheet

A good ML report answers:

1.  What was predicted, and why?
2.  How was the data split and processed?
3.  How was the model tuned?
4.  How did it perform on unseen data?
5.  How should results be interpreted (carefully)?
6.  What are the limitations?

# End-to-end flowchart (zoom-in required)

```{r, eval = TRUE, include = TRUE, echo = FALSE}
grViz("
digraph flow {
  graph [layout = dot, rankdir = LR, size = '12,4!']
  node [shape = box, style = rounded]

  A [label = 'Cleaned and preprocessed data']
  B [label = 'rsplit object']
  C [label = 'training set']
  D [label = 'test set']
  E [label = 'recipe object']
  EMPTY1 [label = 'Specify model']
  F [label = 'parsnip model specification object']
  EMPTY2 [label = 'Start creating workflow object', shape = none]
  G [label = 'workflow object']
  H [label = 'Design resampling scheme']
  I [label = 'rset (resampling set) object']
  J [label = 'Design tuning scheme']
  K [label = 'parameters object']
  L [label = 'tune_grid()', shape = none]
  M [label = 'tune results object']
  N [label = 'best results tibble']
  O [label = 'finalized workflow object']
  P [label = 'last_fit()', shape = none]
  Q [label = 'test set results tibble (final)']

  A -> B [label = 'initial_split()']
  B -> C [label = 'train()']
  B -> D [label = 'test()']
  C -> E [label = 'recipe() %>% step_normalize() %>% \nstep_novel() %>% \nstep_dummy()']
  EMPTY1 -> F [label = 'rand_forest(mtry = tune(), trees = tune()) %>%\nset_engine() %>%\nset_mode()']
  EMPTY2 -> G [label = 'workflow()']
  E -> G [label = 'add_recipe()']
  F -> G [label = 'add_model()']
  H -> I [label = 'vfold_cv()']
  J -> K [label = 'parameter()']
  G -> L [label = '']
  I -> L [label = '']
  K -> L [label = '']
  L -> M [label = '']
  M -> N [label = 'select_best()']
  N -> O [label = 'finalize_workflow()']
  B -> P [label = '']
  O -> P [label = '']
  P -> Q [label = '']

}
")

```
